---
author: "`r Sys.getenv('USERNAME')`"
date: "`r Sys.time()`"
output: 
  html_document:
    code_folding: show
    toc: true
    # toc_float: true
    number_sections: true
    # self_contained: no
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 80
params:
  chr_number: 19
  
  # (root_dir = the chromosome subdirectory)
  # (is dependent on chr_number, so its initiated here, but updated later) 
  root_dir: "S:/Project/SAM/Julia/Ellisif/Paper 1/DATA GWAS/Chr**_parallel_processing_partition"
  
  # Stratification scheme definitions:
  scheme_states: !r c( "A1_TSS1" = "98_TSS1", "A2_TSS2" = "99_TSS2", "Z_znf1" = "84_znf1", "Z_znf2" = "85_znf2" )
  
  # Directory to export resulting files to:
  stage_dir: "0405"

title: "`r paste0('0405 Combine the EWAS probes from all stratification schemes - Chromosome ', params$chr_number) `"
---


```{r}
# Start timer:
start_time <- Sys.time()
```


Combine the stratification scheme-specific files from 0404 into one file, as well as produce tables with overview of number of probes/state loci per scheme and chromosome.


# Make parameters into variables

The pipeline won't be Rmarkdown-based, so make variables containing the parameters stated in the YAML so that the process with building the package later won't be too arduous.

```{r}
chr_number <- params$chr_number

root_dir <- params$root_dir

scheme_states <- params$scheme_states

stage_dir <- params$stage_dir

# Specify path(s) of the scheme-specific file(s) produced in 0404 with the
# scheme-assigned probes :
scheme_info_df <- data.frame(
  scheme = names(scheme_states)
  , state = scheme_states
  , path = file.path( "0404", paste0( "0404_"
                                      , names(scheme_states)
                                      , "_all_scheme_probes.feather" ) )
)

# Specify path for the data frame with the probes from all of the schemes:
all_scheme_probes_path <- file.path( stage_dir
                                  , paste0( stage_dir
                                  , "_all_scheme_probes.feather" ) )
```



# Create stage directory for output

```{r}
if( !dir.exists( stage_dir ) ){
  dir.create( stage_dir )
}

# Delete files (if any) left over from previous runs:
files_in_stage_dir <- list.files( stage_dir, full.names = TRUE )

if( length(files_in_stage_dir) > 0 ){
  message( "Deleting files that were already present in ", stage_dir, "..." )
  lapply( files_in_stage_dir, file.remove )
}
```


# Initial information and set-up

```{r include=FALSE}
# KNIT HOOK THAT ALLOWS FOR FOLDING CHUNK OUTPUT WHEN SPECIFIED
local({
  hooks <-  knitr::knit_hooks$get()
  hook_foldable <- function( type ){
    force(type)
    function( x, options ){
      res <-  hooks[[type]](x, options)

      if( base::isTRUE( options[[paste0( "fold.", type)]])){

        paste0(
          # "\n\n<details><summary>", type, "</summary>\n\n",
          "<details><summary>Click here</summary>\n\n",
          res,
          "\n\n</details>"
        )
      }
      else return(res)
    }
  }

  knitr::knit_hooks$set(
    output = hook_foldable("output"),
    plot = hook_foldable("plot")
  )
})
```

## Session info

Working directory: `r getwd()`\
R_LIBS_USER: `r Sys.getenv('R_LIBS_USER')`\
R_HOME: `r gsub(pattern = "~", replacement = " ~ ", Sys.getenv('R_HOME'))`\
HOME: `r Sys.getenv('HOME')`

```{r fold.output=TRUE}
sessionInfo()
```

## Files used by this script

```{r}
for( i in 1:length( names(params$scheme_states)) ){
  cat( file.path(
    root_dir, "0404",
    paste0( "0404_"
            , names(params$scheme_states)[i]
            , "_all_scheme_probes.feather\n\n" ) )
  )
}
```

## Files produced by this script

-   `r all_scheme_probes_path` -
A data frame containing all probes assigned to a scheme. One file per chromosome.
    

## Rmd set up

```{r setup, include=TRUE}
knitr::opts_chunk$set(
  echo = TRUE
  # , include = FALSE # use to check report text without running any chunk code
  , warning = FALSE # hide warnings
  , message = TRUE # hide messages
  )

options( scipen = 20)
```


## Packages

```{r}
library(dplyr)
library(feather)
library(gt)
library(magrittr)
```

# Import data

## Import annotated probes from all the schemes implemented in 0404

### File metadata

```{r }
metadata <- file.info( 
  list.files(
    path = file.path( root_dir, "0404" )
    , full.names = TRUE
  )
  , extra_cols = FALSE
)
metadata <- cbind( data.frame( "file" = rownames(metadata))
                   , data.frame( metadata, row.names = NULL) ) %>% 
  select( -isdir, -mode, -atime ) %>% 
  filter( grepl( "scheme_probes", file) ) %>% 
  mutate( 
    file = gsub( file.path( root_dir, "0404"), "", file)
  )

metadata %>% 
  gt::gt() %>% 
  fmt_integer( . , columns = 2, use_seps = TRUE) %>% 
  tab_options( table.font.size = "x-small") %>% 
  opt_row_striping()
  
```

### Import files from all schemes with all chromosomes

```{r}
# For all schemes...
probes <- lapply(1:nrow(scheme_info_df), function(s){
  file_name <- scheme_info_df$path[s]
  
  # If the file with this combination of scheme name and chr exists, import
  # from 0404 and place in list:
  if( file.exists(file_name) ){
    feather::read_feather( file_name ) # return imported file
  } else {
    # If not, something internal has gone wrong:
    warning( "file_name = "
             , file_name
             , " does not exist! It will not be included in the combined list "
             , "of all scheme-assigned probes." )
    return()
  }
  
})
length(probes)
sapply(probes, function(x) unique(x$scheme))
sapply(probes, dim)
```


# Combine all the scheme-specific data frames into one large data frame

```{r}
# Combine all data frames into one data frame:
probes <- dplyr::bind_rows( probes )
dim(probes)

probes %>% 
  head() %>% 
  gt::gt() %>% 
  fmt_integer() %>% 
  tab_options( table.font.size = "x-small")
```



# Summary of the data from all the 0404 files

```{r message=FALSE}
probes %>% 
  group_by( scheme, probe_chr) %>% 
  summarise(
    "No of rows" = n()
    , "No of unique schemes present" = n_distinct( scheme )
    , "No of unique EWAS probes" = n_distinct( probe_id )
    , "No of unique combinations of probe ID and scheme" = 
      n_distinct( scheme, probe_id )
    , "No of unique state loci (intervals defined as one feature)" = 
      n_distinct( state_locus )
    , "No of unique combinations of probe ID and state loci" = 
      n_distinct( state_locus, probe_id )
  ) %>%
  ungroup() %>% 
  # Transpose by pivot longer and then pivot wider:
  tidyr::pivot_longer( cols = -c(scheme, probe_chr)
                       , names_to = "stat"
                       , values_to = "n" ) %>% 
  tidyr::pivot_wider( names_from = scheme, values_from = n ) %>% 
  janitor::adorn_totals( where = "col", name = "All schemes" ) %>% 
  # Group by chromosome:
  group_by( probe_chr ) %>% 
  gt::gt( rowname_col = "stat" ) %>%
  tab_options( table.font.size = "x-small") %>%
  tab_options( row_group.background.color = "#fff8f1") %>% 
  fmt_integer( . , columns = -c(1:2), use_seps = TRUE) %>%
  opt_row_striping() %>% 
  gt::tab_header( 
    title =  md( "Summary of the data from 0404, i.e. all the probes that were 
    assigned to a scheme in 0404.  
    Grouped by scheme." )
    , subtitle = md("0404_[scheme_name]_all_scheme_probes.feather  
                    Probes from all stratification schemes combined")
    ) %>% 
  opt_align_table_header( "left" )
```


# Export data frame

Export the large combined data frame with all the EWAS probes assigned to a stratification scheme.

```{r}
message("Exporting ", all_scheme_probes_path, ". . .")
probes %>% 
    feather::write_feather( . , all_scheme_probes_path)

metadata <- file.info( all_scheme_probes_path, extra_cols = FALSE)
metadata <- cbind( data.frame( "file" = rownames(metadata))
                   , data.frame( metadata, row.names = NULL) ) %>% 
  select( -isdir, -mode, -atime ) %>% 
  filter( grepl( "scheme_probes", file) )
metadata %>%
  gt::gt() %>% 
  fmt_integer() %>% 
  tab_options( table.font.size = "x-small")
```



```{r}
# Stop timer:
end_time <- Sys.time()

script_execution_time <- end_time -start_time

cat("The execution time of this script was", as.numeric( script_execution_time, units = "secs" ), "seconds.")
```


# The execution time of this script was __`r round( as.numeric( script_execution_time, units = "mins" ), 3)` minutes__.

# Log execution time

Export data frame with execution time for later collation with execution times of the other stages so that one can create tables with the execution time chromosome and/or stage.

```{r}
# Create folder for csv file if it does not already exist
pipeline_dir <- dirname( dirname( root_dir ) )
stage_execution_time_dir <- 
  file.path( pipeline_dir, "Results", "Preprocessing_stage_execution_times" )

if( !dir.exists( stage_execution_time_dir ) ){
  dir.create( stage_execution_time_dir )
}

exec_time_df <- data.frame(
  stage = stage_dir
  , chr = chr_number
  , script_execution_time_seconds = 
    as.numeric( script_execution_time, units = "secs" )
  , script_start_time = start_time
  , computername = Sys.getenv("COMPUTERNAME")
  , pipeline_dir = pipeline_dir
  , cpu_model = benchmarkme::get_cpu()$model_name
  , no_of_cores = benchmarkme::get_cpu()$no_of_cores
  , ram_iec_units = print(benchmarkme::get_ram(), unit_system = "iec")
  , system_memory_total_Mb = ps::ps_system_memory()$total / 1024^2
  , system_memory_avail_Mb = ps::ps_system_memory()$avail / 1024^2
  , R_platform = R.version$platform
  , R_version = R.version$version.string
  , Platform_GUI = .Platform$GUI
  , RStudio_version = ifelse( .Platform$GUI == "RStudio"
                              , yes = as.character(rstudioapi::getVersion())
                              , no = NA )
)

# Display data frame with log data frame
exec_time_df %>% 
  gt::gt() %>% 
  tab_options( table.font.size = "x-small" ) %>% 
  tab_header( title = md( paste0( "Stage execution time log - Stage "
                                      , stage_dir
                                      , " - Chromosome "
                                      , chr_number )
  ) )

# Export data frame to stage_execution_time_dir
exec_time_df %>% readr::write_csv2( 
  . 
  , file = file.path( stage_execution_time_dir
                 , paste0( stage_dir, "_", sprintf("chr%02d.csv", chr_number )))
  , append = FALSE # overwrite existing files
)
```


# Complete session info

```{r fold.output=TRUE}
sessionInfo()
```

